\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-08-25 analysis/paper/paper.Rmd
switch to plot method for hazard

As we can see, the best fit is the third model. By visualizing the hazard functions for both exposed and unexposed individuals, we can more clearly see how the hazards are no longer proportional.

```{r stanford-hazard, eval = eval_cs4, echo=TRUE, dependson="stanford_fit"}
# Compute hazards---
# First, create a list of time points for both exposure status
hazard_data <- expand.grid(
  exposure = c(0, 1),
  futime = seq(0, 425,
    length.out = 100
  )
)

# Set the offset to zero
hazard_data$offset <- 0
# Use predict to get the fitted values, and exponentiate to
# transform to the right scale
hazard_data$hazard <- exp(predict(fit3,
  newdata = hazard_data,
  type = "link"
))
# Add labels for plots
hazard_data$Status <- factor(hazard_data$exposure,
  labels = c("No Tx", "Tx")
)

ggplot(hazard_data, aes(futime, hazard, colour = Status)) +
  geom_line() +
  paper_gg_theme +  
  ylab("Hazard") +
  xlab("Follow-up time (in days)")
```

@@

```{r stanford-risk, eval = eval_cs4, dependson="stanford_fit"}
# Compute absolute risk curves
newdata <- data.frame(exposure = c(0, 1))
absrisk <- absoluteRisk(fit3,
  newdata = newdata,
  time = seq(0, 425, length.out = 100)
)

@@

plot(absrisk,
     id.names = c("No Tx", "Tx")) + 
  ylab("Cumulative Incidence") +
  xlab("Follow-up time (in days)") +
  paper_gg_theme

# colnames(absrisk) <- c("Time", "NoTrans", "Trans")
# 
# # Rearrange the data
# absrisk <- gather(
#   as.data.frame(absrisk),
#   "Status", "Risk", -Time
# )
# 
# ggplot(absrisk, aes(Time, Risk, colour = Status)) +
#   geom_line() +
#   theme_minimal() +
#   theme(legend.position = "top") +
#   expand_limits(y = 0.5) +
#   xlab("Follow-up time") +
#   ylab("Cum. Incidence")
```

@@

We can also plot the hazard ratio:

```{r, eval = eval_cs4, dependson="stanford_fit"}
newtime <- seq(0, 425, length.out = 100)

>>>>

As we can see, the best fit is the third model. By visualizing the hazard functions for both exposed and unexposed individuals, we can more clearly see how the hazards are no longer proportional. We can easily obtain a plot of the hazards by using the `plot.singleEventCB` method:

```{r stanford-hazard, eval = eval_cs4, echo=TRUE, dependson="stanford_fit", message = FALSE, results="hide"}
plot(fit3, hazard.params = list(xvar = "futime",
                                by = "exposure",
                                alpha = 0.05,
                                ylab = "Hazard",
                                data = cb_data))
```

@@

```{r stanford-risk, eval = eval_cs4, echo = TRUE, dependson="stanford_fit"}
# Compute absolute risk curves
newdata <- data.frame(exposure = c(0, 1))
absrisk <- absoluteRisk(fit3,
  newdata = newdata,
  time = seq(0, 425, length.out = 100)
)

@@

plot(absrisk,
     id.names = c("No Tx", "Tx")) + 
  ylab("Cumulative Incidence") +
  xlab("Follow-up time (in days)") +
  paper_gg_theme
```

@@

<!-- We can also plot the hazard ratio: -->

```{r, eval = FALSE, dependson="stanford_fit"}
newtime <- seq(0, 425, length.out = 100)

### 2020-08-25 analysis/paper/paper.tex
switch to plot method for hazard

As we can see, the best fit is the third model. By visualizing the
hazard functions for both exposed and unexposed individuals, we can more
clearly see how the hazards are no longer proportional.

@@

\begin{CodeInput}
R> # Compute hazards---
R> # First, create a list of time points for both exposure status
R> hazard_data <- expand.grid(
R+   exposure = c(0, 1),
R+   futime = seq(0, 425,
R+     length.out = 100
R+   )
R+ )
R> 
R> # Set the offset to zero
R> hazard_data$offset <- 0
R> # Use predict to get the fitted values, and exponentiate to
R> # transform to the right scale
R> hazard_data$hazard <- exp(predict(fit3,
R+   newdata = hazard_data,
R+   type = "link"
R+ ))
R> # Add labels for plots
R> hazard_data$Status <- factor(hazard_data$exposure,
R+   labels = c("No Tx", "Tx")
R+ )
R> 
R> ggplot(hazard_data, aes(futime, hazard, colour = Status)) +
R+   geom_line() +
R+   paper_gg_theme +  
R+   ylab("Hazard") +
R+   xlab("Follow-up time (in days)")
\end{CodeInput}

@@

\begin{CodeOutput}
#> [1] "absRiskCB" "matrix"
\end{CodeOutput}

@@

We can also plot the hazard ratio:

\begin{center}\includegraphics[width=\textwidth,keepaspectratio=true]{../figures/unnamed-chunk-11-1} \end{center}

@@

This report was generated on 2020-08-21 16:53:49 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.2 (2019-12-12)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Pop!_OS 19.10
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/openblas/libblas.so.3
#> LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.7.so
#> 
#> locale:
#>  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
#>  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
#>  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.7.0               dotwhisker_0.5.0         
#>  [3] timereg_1.9.4             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.0              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.1.0               ggplot2_3.3.2.9000       
#> [13] tidyverse_1.3.0           lubridate_1.7.4          
#> [15] visreg_2.7.0              tibble_3.0.1             
#> [17] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.1-8           
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_4.0-2             
#> [25] Matrix_1.2-18             data.table_1.12.8        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0           
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] ggstance_0.3.3      htmlTable_1.13.3    base64enc_0.1-3    
#>  [7] fs_1.3.2            rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.0-12     
#> [13] xml2_1.3.0          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.7.0      cluster_2.1.0      
#> [19] dbplyr_1.4.2        png_0.1-7           compiler_3.6.2     
#> [22] httr_1.4.1          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.54       tools_3.6.2         gtable_0.3.0       
#> [31] Rcpp_1.0.4.6        cellranger_1.1.0    vctrs_0.3.1        
#> [34] nlme_3.1-143        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.5         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.5       zoo_1.8-7          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.13        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.6          shape_1.4.4        
#> [58] rlang_0.4.6         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-38     labeling_0.3        htmlwidgets_1.5.1  
#> [64] cmprsk_2.2-9        tidyselect_1.1.0    deSolve_1.27.1     
#> [67] plyr_1.8.6          R6_2.4.1            generics_0.0.2     
#> [70] Hmisc_4.3-1         multcomp_1.4-12     DBI_1.1.0          
#> [73] withr_2.2.0         pillar_1.4.4        haven_2.3.1        
#> [76] foreign_0.8-74      mgcv_1.8-31         nnet_7.3-12        
#> [79] mstate_0.2.12       modelr_0.1.5        crayon_1.3.4       
#> [82] rmarkdown_2.3       jpeg_0.1-8.1        grid_3.6.2         
#> [85] readxl_1.3.1        reprex_0.3.0        digest_0.6.25      
#> [88] webshot_0.5.2       numDeriv_2016.8-1.1 stats4_3.6.2       
#> [91] munsell_0.5.0       viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    fix-cs1 /home/sahir/git_repositories/cbpaper
#> Remote:   fix-cs1 @ origin (https://github.com/sahirbhatnagar/cbpaper.git)
#> Head:     [349d153] 2020-08-21: fix merge conflicts
\end{CodeOutput}
\end{CodeChunk}

>>>>

As we can see, the best fit is the third model. By visualizing the
hazard functions for both exposed and unexposed individuals, we can more
clearly see how the hazards are no longer proportional. We can easily
obtain a plot of the hazards by using the \texttt{plot.singleEventCB}
method:

@@

\begin{CodeInput}
R> plot(fit3, hazard.params = list(xvar = "futime",
R+                                 by = "exposure",
R+                                 alpha = 0.05,
R+                                 ylab = "Hazard",
R+                                 data = cb_data))
\end{CodeInput}

@@

\begin{CodeInput}
R> # Compute absolute risk curves
R> newdata <- data.frame(exposure = c(0, 1))
R> absrisk <- absoluteRisk(fit3,
R+   newdata = newdata,
R+   time = seq(0, 425, length.out = 100)
R+ )
R> 
R> class(absrisk)
\end{CodeInput}

\begin{CodeOutput}
#> [1] "absRiskCB" "matrix"    "array"
\end{CodeOutput}

\begin{CodeInput}
R> 
R> plot(absrisk,
R+      id.names = c("No Tx", "Tx")) + 
R+   ylab("Cumulative Incidence") +
R+   xlab("Follow-up time (in days)") +
R+   paper_gg_theme
\end{CodeInput}

@@

@@

This report was generated on 2020-08-25 10:58:13 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 4.0.2 (2020-06-22)
#> Platform: x86_64-apple-darwin17.0 (64-bit)
#> Running under: macOS Catalina 10.15.6
#> 
#> Matrix products: default
#> BLAS:   /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRblas.dylib
#> LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib
#> 
#> locale:
#> [1] en_CA.UTF-8/en_CA.UTF-8/en_CA.UTF-8/C/en_CA.UTF-8/en_CA.UTF-8
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.7.0               dotwhisker_0.5.0         
#>  [3] timereg_1.9.4             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.2              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.1.1               ggplot2_3.3.2            
#> [13] tidyverse_1.3.0           lubridate_1.7.9          
#> [15] visreg_2.7.0              tibble_3.0.3             
#> [17] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.2-3           
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_4.0-2             
#> [25] Matrix_1.2-18             data.table_1.13.0        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0.9000      
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] ggstance_0.3.4      htmlTable_1.13.3    base64enc_0.1-3    
#>  [7] fs_1.4.1            rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.1-1      
#> [13] xml2_1.3.2          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.7.0      cluster_2.1.0      
#> [19] dbplyr_1.4.4        png_0.1-7           compiler_4.0.2     
#> [22] httr_1.4.2          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.55       tools_4.0.2         gtable_0.3.0       
#> [31] Rcpp_1.0.5          cellranger_1.1.0    vctrs_0.3.2        
#> [34] nlme_3.1-148        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.6         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.6       zoo_1.8-8          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.14        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.7          shape_1.4.4        
#> [58] rlang_0.4.7         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-41     labeling_0.3        htmlwidgets_1.5.1  
#> [64] cmprsk_2.2-10       tidyselect_1.1.0    deSolve_1.28       
#> [67] R6_2.4.1            generics_0.0.2      Hmisc_4.4-0        
#> [70] multcomp_1.4-13     DBI_1.1.0           withr_2.2.0        
#> [73] pillar_1.4.6        haven_2.3.1         foreign_0.8-80     
#> [76] mgcv_1.8-31         nnet_7.3-14         mstate_0.2.12      
#> [79] modelr_0.1.8        crayon_1.3.4        rmarkdown_2.3      
#> [82] jpeg_0.1-8.1        grid_4.0.2          readxl_1.3.1       
#> [85] blob_1.2.1          reprex_0.3.0        digest_0.6.25      
#> [88] webshot_0.5.2       numDeriv_2016.8-1.1 stats4_4.0.2       
#> [91] munsell_0.5.0       viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    editMax /Users/maxturgeon/Documents/git_repos/cbpaper
#> Head:     [08c024e] 2020-08-21: Merge pull request #26 from sahirbhatnagar/fix-cs1
\end{CodeOutput}
\end{CodeChunk}
